# Multiple Variable Analisys Exercise

## Working with diamonds dataset

### Diamonds price histogram by color

```{r}
library('dplyr')
library('tidyr')
library(ggplot2)

data(diamonds)
ggplot(aes(x = price, fill = cut), data = diamonds)+
  facet_wrap( ~ color )+
  scale_x_log10()+
  geom_histogram()+
  scale_fill_brewer(type = 'qual')
```

### Price vs Table Scatter

```{r}
ggplot(aes(x = table, y = price,color = cut), data = diamonds)+
  geom_point()+
  scale_x_continuous(limits= c(50,80), breaks = seq(50, 80, 2))+
  scale_color_brewer(type = 'qual')
```

### Diamond volume vs price

```{r}
diamonds$volume <- diamonds$x*diamonds$y*diamonds$z
summary(diamonds$volume)
ggplot(aes(y=price, x=volume, color=clarity), 
       data=subset(diamonds, volume<=quantile(volume, 0.99)))+
  geom_point()+
  scale_y_log10()+
  scale_color_brewer(type = 'div')
```

## Facebook dataset

### Friendship initiated vs Tenure

```{r}
pf <- read.delim('pseudo_facebook.tsv')
str(pf)
pf$prop_initiated <- pf$friendships_initiated/pf$friend_count
summary(pf$prop_initiated)
```
```{r}
pf$year_joined <- floor(2014 - pf$tenure / 365)
pf$year_joined.bucket <- cut(x=pf$year_joined, breaks = c(2004,2009,2011,2012,2014))
summary(pf$tenure)
table(pf$prop_initiated)

ggplot(aes(y = prop_initiated, x=tenure, color=year_joined.bucket), 
       data=subset(pf, !is.na(prop_initiated) & !is.na(year_joined.bucket)))+
  geom_line(stat = 'summary', fun.y = median)
```

### Friendship initiated vs Tenure - Smooth

```{r}
ggplot(aes(y = prop_initiated, x=tenure, color=year_joined.bucket), 
       data=subset(pf, !is.na(prop_initiated) & !is.na(year_joined.bucket)))+
  geom_smooth()
```

```{r}
by(pf$prop_initiated, pf$year_joined.bucket, summary)
```

## Back to diamonds dataset

### Price/carat ratio

```{r, results=asis}
ggplot(aes(y=price/carat, x=cut, color=color), data=diamonds)+
  facet_wrap( ~clarity )+
  geom_jitter()+ #also works with: geom_point(position = position_jitter(h=0))+
  scale_color_brewer(type = 'div')

```

## Health Risk Dataset

### Preparing dataframe

```{r}
library(xlsx)
library('tidyr')
library('dplyr')

dados <- array(c('Indicator_TC male ASM.xlsx','M','TC',
                 'Indicator_TC female ASM.xlsx','F','TC',
                 'Indicator_BMI male ASM.xlsx','M','BMI',
                 'Indicator_BMI female ASM.xlsx','F','BMI',
                 'Indicator_SBP male ASM.xlsx','M','SBP',
                 'Indicator_SBP female ASM.xlsx','F','SBP'), dim=c(3,6,2))

print(dados)

#health_risk <- read.xlsx('Indicator_TC male ASM.xlsx', sheetIndex = 1, check.names = F)
#health_risk$gender <- 'M'
#health_risk$indicator <- 'TC'
#names(health_risk)[1] <- "country"

rm(health_risk)
rm(health_risk_tmp)

for (i in 1:6) {
  print(dados[3*(i-1)+1])
  print(dados[3*(i-1)+2])
  print(dados[3*(i-1)+3])
  
  
  if (!exists("health_risk")) {
    
    health_risk <- read.xlsx(dados[3*(i-1)+1], sheetIndex = 1, check.names = F)
    health_risk$gender <- dados[3*(i-1)+2]
    health_risk$indicator <- dados[3*(i-1)+3]
    names(health_risk)[1] <- "country"
    
  } else {
    health_risk_tmp <- read.xlsx(dados[3*(i-1)+1], sheetIndex = 1, check.names = F)
    health_risk_tmp$gender <- dados[3*(i-1)+2]
    health_risk_tmp$indicator <- dados[3*(i-1)+3]
    names(health_risk_tmp)[1] <- "country"
    
    health_risk <- bind_rows(health_risk, health_risk_tmp)
    
  }
  
}

rm(health_risk_tmp)

health_risk <- gather(health_risk, "year", "value", 2:30)
health_risk$gender <- factor(health_risk$gender)
health_risk$indicator <- factor(health_risk$indicator)
```

### Plots

```{r}
ggplot(aes(x=year, y=value, color=gender), 
       data=subset(health_risk, indicator == "SBP"))+
  geom_point()+
  facet_wrap( ~indicator, nrow=3)
```

## Water use dataset

### Preparing data

```{r}

water <- read.csv('urban_water.csv', check.names = FALSE)
water$local <- 'urban'
names(water)[1] <- "country"
water_tmp <- read.csv('rural_water.csv', check.names = FALSE)
water_tmp$local <- 'rural'
names(water_tmp)[1] <- "country"

water <- bind_rows(water, water_tmp)

rm(water_tmp)

names(water)

water <- gather(water, "year", "value", 2:22)
water <- subset(water, !is.na(value))

water$local <- factor(water$local)
water$year <- as.numeric(water$year)

water$continent <- NA


europe = c("Norway", "France", "Portugal", "Finland", "Germany", 
           "Denmark", "Spain", "Ireland", "Iceland", "Switzerland",
           "Netherlands", "Ukraine", "Sweden", "Russia", "United Kingdom",
           "Slovenia", "Hungary", "Romania", "Austria", "Albania", "Italy",
           "Belgium", "Bulgaria", "Luxembourg", "Greece", "Croatia",
           "Czech Republic", "Motenegro", "Malta", "Lithuania", "Liechtenstein",
           "Poland", "Slovak Republic", "Moldova", "Turkey", "Latvia",
           "Serbia", "Armenia", "Estonia", "Montenegro", "Macedonia, FYR",
           "Bosnia and Herzegovina", "Belarus", "Georgia","Andorra","Monaco","San Marino","Holy See")
water$continent <- ifelse(water$country %in% europe, "Europe", water$continent)

africa = c("Niger", "Mozambique", "Congo, Rep.", "Ghana", "Sudan",
           "South Africa", "Madagascar", "Angola", "Algeria", "Uganda",
           "Rwanda", "Sierra Leone", "Ethiopia", "Central African Republic",
           "Burkina Faso", "Zimbabwe", "Chad", "Liberia", "Burundi",
           "Cote d'Ivoire", "Guinea-Bissau", "Guinea", "Malawi", "Tanzania",
           "Mali","Senegal", "Togo", "Morocco", "Cameroon", "Cape Verde",
           "Botswana", "Benin", "Congo, Dem. Rep.", "Egypt", "Gabon",
           "Mauritania", "Zambia", "Tunisia", "Kenya", "Cambodia",
           "Nigeria", "Comoros", "Eritrea", "Mauritius", "Seychelles",
           "Gambia","Lesotho","Libya", "Sao Tome and Principe","Swaziland",
           "Djibouti","Equatorial Guinea", "Somalia","Ghana")
water$continent <- ifelse(water$country %in% africa, "Africa", water$continent)

southamerica = c("Brazil", "Argentina", "Uruguay", "Chile", "Bolivia",
                 "Paraguay", "Venezuela", "Peru", "Colombia", "Guyana",
                 "Ecuador", "Suriname")
water$continent <- ifelse(water$country %in% southamerica, "South America", water$continent)

northamerica = c("United States", "Canada", "Greenland")
water$continent <- ifelse(water$country %in% northamerica, "North America", water$continent)

asia = c("Japan", "China", "Thailand", "India", "Bangladesh", "North Korea",
         "Saudi Arabia", "Afghanistan", "Qatar", "South Korea", "Bahrain",
         "Iran", "Israel", "Brunei", "Philippines", "Syria",
         "Mongolia", "Namibia", "Pakistan", "Lao", "Maldives",
         "Iraq", "Yemen", "Sri Lanka", "United Arab Emirates", "Tajikistan",
         "Lebanon", "Jordan", "Kuwait", "Oman", "Singapore",
         "Hong Kong, China", "Malaysia", "Nepal", "Vietnam", "Indonesia",
         "Myanmar", "Cyprus", "Timor-Leste", "Uzbekistan", "Turkmenistan",
         "Kyrgyz Republic", "Kazakhstan", "Bhutan", "Azerbaijan", "Macao, China")
water$continent <- ifelse(water$country %in% asia, "Asia", water$continent)

centralamerica = c("Honduras", "Costa Rica", "Mexico", "Panama", "Bahamas",
                   "Haiti", "Jamaica", "Cuba", "Dominican Republic", "Belize",
                   "El Salvador", "Guatemala", "Barbados", "Dominica", 
                   "Trinidad and Tobago","Nicaragua","Antigua and Barbuda",
                   "St. Vincent and the Grenadines","St. Kitts and Nevis",
                   "St. Lucia", "Grenada","Bermuda","Aruba")
water$continent <- ifelse(water$country %in% centralamerica, "Central America", water$continent)

oceania = c("Australia", "New Zealand", "Fiji", "Tonga", "Palau",
            "Solomon Islands", "Palau", "Samoa", 
            "Papua New Guinea", "Micronesia, Fed. Sts.",
            "Vanuatu", "Kiribati")
water$continent <- ifelse(water$country %in% oceania, "Oceania", water$continent)
```

### Plots

```{r}
ggplot(aes(x=year, y=value, color = continent), 
       data=subset(water, !is.na(water$continent)))+
  geom_line(stat = 'summary', fun.y = 'min')+
  facet_wrap( ~local ) 
```

```{r}
ggplot(aes(x=year, y=value, color = local), data=water)+
  geom_jitter()+
  facet_wrap( ~continent )
```

```{r}
ggplot(aes(x=local, y=value, color = continent), data=water)+
  geom_jitter()
```
